---
title: Agricultural land vs. agriculural value added to GDP
author: Anicet Ebou
date: '2020-01-05'
slug: agricultural-land-vs-agriculural-value-added-to-gdp
categories:
  - R
tags:
  - rstats
  - Data Science
  - Agriculture
toc: no
images: ~
---

It comes as an idea to explore different relationship between some variables in many fields. The first one that attract my attention was the relation between agricultural land (% of land area) and value added by agricuture to GDP. For comparison I choose West African countries (Cote d'Ivoire, Ghana, Nigeria) and some other countries (USA, Japan, UK). The data for this analysis come from [gapminder](https://gapminder.org/data).

## Data import and cleaning

The raw data is part of the [afrods](https://github.com/Ebedthan/afrods.git) package and can be downloaded [here](https://github.com/Ebedthan/afrods/tree/master/data-raw/xls/Agriculture%20and%20Environment)(the two files are Agriculture (p of GDP).xlsx and agriculture land.xlsx). The files report data on the period from 1960 to 2010.

To rapidly tidy up the data raw I use the [tidygapminder](https://github.com/Ebedthan/tidygapminder.git) package.

First we import the tidygapminder and tidyverse package.

```{r warning=FALSE, message=FALSE}
library(tidygapminder)
library(tidyverse)
```

Second, you need to put into tidy the two .xls files and merge them into one big data frame using the tidygapminder `tidy_bunch` function.

```{r echo=FALSE, message=FALSE}
# Create first table for demo
agri_data <- tidy_bunch(dirpath = "/home/ediman/Projects/2019_workshop-agence-emploi-jeunes/02-d1-transform-viz/data/agri", merge = TRUE)
```

```
# Create first table for demo
agri_data <- tidy_bunch(dirpath = "path/to/data/folder", merge = TRUE)
```

We can now change some variable name to easily manipulate them.

```{r}
agri_data <- agri_data %>% 
  rename(agri_val_added = `Agriculture, value added (% of GDP)`, land_area = `Agricultural land (% of land area)`) %>% filter(country %in% c("Cote d'Ivoire", "Ghana", "Nigeria", "Japan", "United States", "France"))
```

# Data exploration

We can now explore the data to see any relationship

```{r}
ggplot(data = agri_data, mapping = aes(x = land_area, y = agri_val_added, color = country)) + geom_point() + 
  labs(title = "Agriculture value added to GDP vs. Agricultural land", 
       x = "agricultural land (%)", y = "Agriculture contribution to GDP (%)")
```


It is pretty interesting to already note that countries like France, USA or Japan have a agricultural land percentage less than 60 % and a contribution to GDP less than 10% while African selected countries have at least 49 % and a contribution to GDP of a least 21 %.

Let us plot this relationship differently.

```{r message=FALSE}
ggplot(data = agri_data, mapping = aes(x = land_area, y = agri_val_added, color = country)) + geom_point() + facet_wrap(~country)
  labs(title = "Agriculture value added to GDP vs. Agricultural land", 
       x = "agricultural land (%)", y = "Agriculture contribution to GDP (%)")
```

We can see that African countries show a common pattern while others countries show another common pattern.

## What could be the explanation ?

One of the possible hypothesis could be the diversification of fields adding value to GDP in countries enabling Japan, USA, or France to have mor value added by industries, fashion industries, etc. while West African countries can be more dependant of agriculture.

To verify this hypothesis we can explore for example the value added by industry to GDP.

For that we take the Industry value added (% to GDP) file from the same source.

```{r}

```