---
title: How to export high quality images in R
author: Anicet Ebou
date: '2020-11-17'
slug: how-to-export-high-quality-images-in-r
categories:
  - R
tags:
  - R
  - rstats
images: ~
---

There is multiple way to export images in R, but the one I prefer is with the package Cairo. Cairo, gives controls on the properties of the final image and such control is always appreciated especially if you're planning to submit images in a scientific publication like it's often the case for me.

First if you haven't yet installed Cairo, install it from CRAN with:

```{r, eval=FALSE}
install.packages('Cairo')
```
  
After that let us create a plot for our example:
```{r}
library(ggplot2)

p <- ggplot(data = diamonds, aes(x = price, y = carat, colour = color)) +
      geom_point()
```

Now we define our properties for the image we will export:

```{r}
Cairo::Cairo(
  13, 8,               # width, height in inch
  units = "in",        # unit of width and height of image, inches
  "diamond_plot.tiff", # name of the exported image
  bg = "white",        # white background
  type = "tiff",       # tiff image file
  dpi = 300            # DPI resolution
)
p
invisible(dev.off())   # dev.off() close the file
```

Here I use one of my favorite trick: the width and height are numbers of the [Fibonacci numbers](https://en.wikipedia.org/wiki/Fibonacci_number). One of the particularity of the Fibonacci numbers is that the ratio of two successive numbers for n greater or equal to 3 is approximate the golden number. So they give a great ratio to images!

For example the beginning of the numbers are:

`1 + 1 + 2 + 3 + 5 + 8 + 13 + 21 + ...`

So for example 2/3 = `r 2/3`.

It gives a great ratio for images and using units like inches ease the use of Fibonacci numbers.

Now we have our image exported successfully!

Hope you've enjoyed and that you'll use it. Don't hesitate to leave a comment if you found it useful.
